### 2009-10-03 paper/paper.tex
Up!

\documentclass{letter}[11pt]

\begin{document}
\begin{letter}{

Test stuff
\end{document}